---
title: "Adegenet From VCF"
output:
  html_document:
    df_print: paged
---

## Load libraries and set working directory
```{r load-libs, message=FALSE, warning=FALSE}
library(adegenet)
library(adegraphics)
library(vcfR)
library(pegas)
library(hierfstat)
library(ggplot2)
library(viridis)
library(tidyverse)

knitr::opts_chunk$set(fig.path = "plots_LD/", dev = c("png", "svg"), dpi=500, cache=T)
```


## Import data and define working object
```{r import-data, message=FALSE, warning=FALSE, results=FALSE, cache=TRUE}

# Import data

vcfFile <- "/mmfs1/scratch/CaribouProject/data/vcf/Caribou.filtered.vcf"
allPop.vcf <- read.vcf(vcfFile, to=30000)

# Transform vcf to genind
allPop.genind <- loci2genind(allPop.vcf)

# Import population ID
popID.df <- read.csv("/mmfs1/scratch/CaribouProject/labels.csv", header = T)[,-1]

#popID.df[popID.df == "Togiak"] <- 'Outgroup'
allPop.genind@pop <- as.factor(popID.df[,2])

# Subset without Togiak
noTog.genind <- allPop.genind[pop=c("West", "Central", "East", "Unk")]

# Copy genind object
brucellosis.genind <- allPop.genind

# Redefine populations based on Brucellosis status
brucellosis.genind@pop <- as.factor(popID.df[,3])
```

## Start analysis

### 	Bartlett test of homogeneity of variances
```{r bartlett-test, cache=TRUE}
allPop.summary <- summary(allPop.genind)

bartlett.test(list(allPop.summary$Hexp,allPop.summary$Hobs))
t.test(allPop.summary$Hexp, allPop.summary$Hobs,pair=T,var.equal=TRUE,alter="greater")

```

### F-stats 
#### Test of Hardy–Weinberg Equilibrium
```{r hwt, message=FALSE, warning=FALSE, cache=TRUE}
allPop.hwt <- hw.test(allPop.genind, B=0)
colMeans(allPop.hwt)
```

##### No Togiak
```{r hwt-ntg, message=FALSE, warning=FALSE, cache=TRUE}
noTog.hwt <- hw.test(noTog.genind, B=0)
colMeans(noTog.hwt)
```

#### General Fst & Fis 
```{r fstats, cache=TRUE}
allPop.fstats <- wc(allPop.genind)
print(paste("FIS:", allPop.fstats$FIS))
print(paste("FST:", allPop.fstats$FST))

## == ##
allPop.basics <- basic.stats(allPop.genind)
allPop.basics$overall
```

##### No Togiak
```{r fstats-ntg}
noTog.fstats <- wc(noTog.genind)
print(paste("FIS:", noTog.fstats$FIS))
print(paste("FST:", noTog.fstats$FST))

## == ##
noTog.basics <- basic.stats(noTog.genind)
noTog.basics$overall
```

#### Bootstraps
```{r bootstraps, cache=TRUE}
betas(allPop.genind, nboot=102, lim=c(0.025,0.975), diploid=TRUE, betaijT=FALSE)
betas(noTog.genind, nboot=102, lim=c(0.025,0.975), diploid=TRUE, betaijT=FALSE)
```


#### General Confidence Interval

```{r general-cv, cache=TRUE}
nc <- genind2hierfstat(allPop.genind)
allPop.CV <- boot.vc(nc[1], nc[-1])$ci

print(allPop.CV)
```

##### No Togiak
```{r general-cv-ntg, cache=TRUE}
nc <- genind2hierfstat(noTog.genind)
noTog.CV <- boot.vc(nc[1], nc[-1])$ci

print(noTog.CV)
```


```{r general-cv-sub, cache=TRUE, eval=FALSE}
ncs <- genind2hierfstat(subPop.genind)
#subPop.CV <- boot.ppfis(nc[1], nc[-1])$ci
subPop.CI <- boot.ppfis(dat=ncs,nboot=100,quant=c(0.025,0.975),diploid=TRUE,dig=4)
print(subPop.CI)
```

##### Per population
```{r fis-pop, cache=TRUE}
for (pop in unique(popID.df[,2])) {
  print(pop)
  
  subPop.genind <- allPop.genind[pop=c(pop)]
  subPop.fstats <- wc(subPop.genind)
  
  print(paste("FIS:", subPop.fstats$FIS))
}
```

### Pairwise Fst
```{r pairwise-fst, cache=TRUE}
matFst <- genet.dist(allPop.genind, method = "Nei87")

matFst %>% 
  as.matrix() %>% 
  as_tibble(rownames = "A") %>% pivot_longer(-A, names_to = "B", values_to = "distance") %>% 
  mutate(distance = na_if(distance, 0),
         A = factor(A, levels = c("West", "Central", "East", "Outgroup")),
         B = factor(B, levels = c("West", "Central", "East", "Outgroup"))) %>% 
  ggplot(aes(x=A, y=B, fill=distance)) +
  geom_tile() + scale_fill_gradient(low = "#e86100", high = "#92bbff") +
  geom_text(aes(label = round(distance, 4))) +
  labs(x = "Population") +
  theme(axis.title.y = element_blank(), legend.position = "right")
```
### Inbreeding

```{r inbreeding, message=FALSE, warning=FALSE, cache=TRUE}
allPop.inbr <- inbreeding(allPop.genind, N = 100)

allPop.FIS <- sapply(allPop.inbr, mean)

hist(allPop.FIS, col="orange", xlab="mean value of F",
     main="Distribution of mean F across individuals")

#No Togiak
noTog.inbr <- inbreeding(noTog.genind, N = 100)

noTog.FIS <- sapply(noTog.inbr, mean)

hist(noTog.FIS, col="blue", xlab="mean value of F",
     main="Distribution of mean F across individuals")

```

### PCA
#### All pop
```{r pca-allpop, cache=TRUE}
# Transform genind object
allPop.gen <- tab(allPop.genind, NA.method = "mean")

# Plot PCA
allPop.pca <- dudi.pca(allPop.gen, scannf=F, nf=15)
#screeplot(allPop.pca, npcs=15)
#summary(allPop.pca)

var.explained.pca <- round(allPop.pca$eig/sum(allPop.pca$eig)*100, digits=1)

## Ellipses are 1.5x intertia for the group
s.class(allPop.pca$li, xax=1, yax=2, col=funky(7), fac=pop(allPop.genind), 
        paxes.draw=T, main="PCA MCH + Outgroup", plabels.cex = F,
        xlab=paste0("PC 1 (", var.explained.pca[1], "%)"), ylab=paste0("PC 2 (", var.explained.pca[2], "%)"))
```

#### No Togiak
```{r pac-notog, cache=TRUE}
# Transform genind object
noTog.gen <- tab(noTog.genind, NA.method = "mean")

# Plot PCA
noTog.pca <- dudi.pca(noTog.gen, scannf=F, nf=15)
#screeplot(allPop.pca, npcs=15)
#summary(allPop.pca)

var.noTog.pca <- round(noTog.pca$eig/sum(noTog.pca$eig)*100, digits=1)

## Ellipses are 1.5x intertia for the group
s.class(noTog.pca$li, xax=1, yax=2, col=funky(7), fac=pop(noTog.genind), 
        paxes.draw=T, main="PCA MCH", plabels.cex = F,
        xlab=paste0("PC 1 (", var.noTog.pca[1], "%)"), ylab=paste0("PC 2 (", var.noTog.pca[2], "%)"))
```

### PCoA
#### All pop
```{r pcoa-allpop, cache=TRUE}
# Make PCoA object
allPop.pcoa <- dudi.pco(dist(allPop.genind), scannf=FALSE, nf=15)

# Plot PCoA
var.explained.pcoa <- round(allPop.pcoa$eig/sum(allPop.pcoa$eig)*100, digits=1)

## Ellipses are 1.5x intertia for the group
s.class(allPop.pcoa$li, xax=1, yax=2, col=funky(7), fac=pop(allPop.genind),
        paxes.draw=T, main="PCoA MCH + Outgroup", plabels.cex = F,
        xlab=paste0("PC 1 (", var.explained.pcoa[1], "%)"), ylab=paste0("PC 2 (", var.explained.pcoa[2], "%)"))

```

```{r, eval=FALSE, cache=TRUE}
rownames(allPop.pcoa$li)[allPop.pcoa$li$A1 > 30]
rownames(allPop.pcoa$li)[allPop.pcoa$li$A1 < -50]
rownames(allPop.pcoa$li)[allPop.pcoa$li$A2 > 50]
rownames(allPop.pcoa$li)[allPop.pcoa$li$A2 < -7]
```

#### No Togiak
```{r pcoa-notog, cache=TRUE}

# Make PCoA object
noTog.pcoa <- dudi.pco(dist(noTog.genind), scannf=FALSE, nf=15)

# Plot PCoA
var.noTog.pcoa <- round(noTog.pcoa$eig/sum(noTog.pcoa$eig)*100, digits=1)

## Ellipses are 1.5x intertia for the group
s.class(noTog.pcoa$li, xax=1, yax=2, col=funky(7), fac=pop(noTog.genind),
        paxes.draw=T, main="PCoA MCH", plabels.cex = F,
        xlab=paste0("PC 1 (", var.noTog.pcoa[1], "%)"), ylab=paste0("PC 2 (", var.noTog.pcoa[2], "%)"))

```

```{r, eval=FALSE}
rownames(noTog.pcoa$li)[noTog.pcoa$li$A1 > 50]
rownames(noTog.pcoa$li)[noTog.pcoa$li$A1 < -50]
rownames(noTog.pcoa$li)[noTog.pcoa$li$A2 > 50]
rownames(noTog.pcoa$li)[noTog.pcoa$li$A2 < -50]
```

### DAPC
#### All pop
```{r dapc-allpop-1}
# Make DAPC object
allPop.dapc <- dapc(allPop.genind, allPop.genind$pop, n.pca = 15, n.da = 15)

# Plot DAPC scatter
scatter(allPop.dapc, legend=TRUE)
```


```{r dapc-allpop-2, fig.height = 5, fig.width = 20, fig.align = "center"}
# Plot DAPC compoplot
compoplot(allPop.dapc, posi="bottomright", show.lab = T, lab = allPop.genind$pop)
```

#### No Togiak
```{r dapc-notog-1}
# Make DAPC object
noTog.dapc <- dapc(noTog.genind, noTog.genind$pop, n.pca = 15, n.da = 15)

# Plot DAPC scatter
scatter(noTog.dapc, legend=TRUE)
```

```{r dapc-notog-2, fig.height = 5, fig.width = 20, fig.align = "center"}
# Plot DAPC compoplot
compoplot(noTog.dapc, posi="bottomright", show.lab = T, lab = noTog.genind$pop)
```

## Brucellosis

### Set data
```{r set-brucellosis, message=FALSE, warning=FALSE, cache=TRUE}
# Copy genind object

brucellosis.genind <- allPop.genind

# Redefine populations based on Brucellosis status
brucellosis.genind@pop <- as.factor(popID.df[,3])

# Remove untested individuals
brucellosis.genind <- brucellosis.genind[pop=c("Neg", "Pos"),]
```

### Pairwise Heterozygosity test
```{r brucellosis-pairwise-het, fig.align = "center", cache=TRUE}
brucellosis.Hs.test <- Hs.test(brucellosis.genind[pop="Neg"],
                brucellosis.genind[pop="Pos"],
                n.sim=499)


brucellosis.Hs.test

plot(brucellosis.Hs.test,main=paste("p.value = ", round(brucellosis.Hs.test$pvalue, dig = 5)))
```

### F-stats
```{r brucellosis-Fstats, fig.align = "center", results="hold", cache=TRUE}

basic <- basic.stats(brucellosis.genind, diploid = TRUE, digits = 2)

mean.Hs <- colMeans(basic$Hs)
mean.Ho <- colMeans(basic$Ho)

print(paste("Expected Heterozigosity Neg:", round(mean.Hs["Neg"], digits = 5)))
print(paste("Expected Heterozigosity Pos:", round(mean.Hs["Pos"], digits = 5)))
print(paste("Observed Heterozigosity Neg:", round(mean.Ho["Neg"], digits = 5)))
print(paste("Observed Heterozigosity Pos:", round(mean.Ho["Pos"], digits = 5)))
```

```
### PCOA
```{r PCOA-brucellosis, fig.align = "center", cache=TRUE}
# Make PCoA object
brucellosis.pcoa <- dudi.pco(dist(brucellosis.genind), scannf=FALSE, nf=15)

# Plot PCoA
var.brucellosis.pcoa <- round(brucellosis.pcoa$eig/sum(brucellosis.pcoa$eig)*100, digits=1)

## Ellipses are 1.5x intertia for the group
s.class(brucellosis.pcoa$li, xax=1, yax=2, col=funky(7), fac=pop(brucellosis.genind),
        paxes.draw=T, main="PCoA Brucellosis", plabels.cex = F,
        xlab=paste0("PC 1 (", var.brucellosis.pcoa[1], "%)"), ylab=paste0("PC 2 (", var.brucellosis.pcoa[2], "%)"))
```







